---
title: "AI Project"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

Here we explore all our available data files

# Set up

```{r set-up, message = FALSE}
library(tidyverse)
library(readxl)
library(gtsummary)

# library(rio)
# library(lubridate)
# library(stringr)
# library(janitor)
# library(ggplot2)
# library(plotly)
# library(purrr)
# library(reactable)
# library(ggvenn)
# library(naniar)
# library(haven)
```

# Data

## Survey

```{r data}
francois_clean_rating_data <- read_excel("data/AI_messages_Eval_6-8-26.xlsx", sheet = "Sheet1")
nutr_rater_data <- read_excel("data/AI_messages_Eval_6-8-26.xlsx", sheet = "Nutr_Eval_DATA_LABELS_2026-04-0")[, 1:11]
PA_rater_data <- read_excel("data/AI_messages_Eval_6-8-26.xlsx", sheet = "PA_eval_DATA_Labels")[, 1:11]
```

## Basic Check

```{r}
francois_clean_rating_data_basic_check <- francois_clean_rating_data %>%
  mutate(
    check1 = case_when(
      `perceived_source (1=human, 2=AI, 3=not sure)` == "3" ~ NA_real_,
      `true_source (1=human, 2=AI)` == `perceived_source (1=human, 2=AI, 3=not sure)` ~ 1,
      `true_source (1=human, 2=AI)` != `perceived_source (1=human, 2=AI, 3=not sure)` ~ 0
    )
  )

identical(francois_clean_rating_data_basic_check$check1, francois_clean_rating_data_basic_check$source_correct_binary) # TRUE
```

## Clean Survey Data

Rename, recode, and factor some variables for clarity

```{r}
clean_rating_data <- francois_clean_rating_data %>%
  mutate(
    domain = recode(domain, "Physiscal activity" = "Physical Activity"),
    domain = recode(domain, "nutrition" = "Nutrition"),
    domain = factor(domain, levels = c("Nutrition", "Physical Activity")),
    true_source = case_when(
      `true_source (1=human, 2=AI)` == 1 ~ "human",
      `true_source (1=human, 2=AI)` == 2 ~ "AI",
      TRUE ~ NA_character_
    ),
    true_source = factor(true_source, levels = c("human", "AI")),
    source_correct_3cat = factor(source_correct_3cat)
  )

str(clean_rating_data)
unique(clean_rating_data$domain)
unique(clean_rating_data$true_source)
```

## Clean Rater Data

```{r}
clean_nutr_rater_data <- nutr_rater_data %>%
  mutate(rater_id = paste0("ne", `Record ID`), domain = "Nutrition") %>%
  dplyr::rename(
    record_id = "Record ID",
    redcap_survey_identifier = "Survey Identifier",
    survey_timestamp = "Survey Timestamp",
    first_name = "First Name",
    last_name = "Last Name",
    credentials = "Highest Degree (e.g. PhD, MD)",
    rank = "Rank (e.g. Postdoc, Instructor, Assistant Professor, etc)",
    years_at_rank = "Years at Rank",
    affiliation = "Institution",
    certifications = "License, Certifications, Professional Certifications (e.g. RDN, ACSM)",
    area_of_expertise = "Area of Research Expertise"
  ) %>%
  mutate(rank = recode(rank, "Associate Profesor" = "Associate Professor")) %>%
  select(
    domain,
    rater_id,
    credentials,
    rank,
    years_at_rank,
    certifications,
    area_of_expertise
  )

clean_PA_rater_data <- PA_rater_data %>%
  mutate(
    rater_id = paste0("pae", record_id),
    certifications = case_when(
      is.na(certifications) ~ NA,
      certifications %in% c("none", "None", "NA", "Na", "na", "N/A", "n/a") ~ NA,
      TRUE ~ certifications
    ), 
    domain = "Physical Activity"
  ) %>%
  dplyr::rename(
    survey_timestamp = "evaluation_of_physical_activity_intervention_messa_timestamp",
    first_name = "first_name_v2",
    last_name = "last_name_v2",
    credentials = "credentials_v2",
    affiliation = "affiliation_v2",
    area_of_expertise = "research_expertise"
  ) %>%
  select(
    domain,
    rater_id,
    credentials,
    rank,
    years_at_rank,
    certifications,
    area_of_expertise
  )

identical(names(clean_nutr_rater_data), names(clean_PA_rater_data))

rater_data <- rbind(clean_nutr_rater_data, clean_PA_rater_data) %>%
  distinct(rater_id, .keep_all = TRUE) %>%
  mutate(
    across(
      c(credentials, rank, certifications, area_of_expertise),
      ~ as.factor(.x)
    ),
    years_at_rank = as.numeric(years_at_rank)
  )

str(rater_data)
```

# Descriptive Analyses

## Message-Level Characteristics

-   Report by domain and true source:

-   number of messages

-   number human-generated

-   number AI-generated

```{r}
# Create one-row-per-message dataset
message_level_df <- clean_rating_data %>%
  select(message_id, domain, true_source) %>%
  distinct()

# Table 1: Message characteristics
table1_messages <- message_level_df %>%
  tbl_summary(
    by = true_source,
    include = domain,
    label = list(
      domain ~ "Domain"
    ),
    statistic = list(
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_categorical() ~ c(0, 1),
    missing = "no"
  ) %>%
  add_overall(last = FALSE) %>%
  modify_header(
    label ~ "",
    stat_0 ~ "**Overall**<br>N = {N}",
    stat_1 ~ "**Human-generated**<br>N = {n}",
    stat_2 ~ "**AI-generated**<br>N = {n}"
  ) %>%
  modify_caption("**Table 1. Messages by domain and true source**") %>%
  bold_labels()

table1_messages
```

## Rater-Level Characteristics

-   Report overall and by domain, if relevant:

-   number of raters

-   credentials

-   rank

-   years of rank

-   certifications

-   area of expertise

```{r}
# Table 1: Rater characteristics
table1_raters_by_domain <- rater_data %>%
  select(
    domain,
    credentials,
    rank,
    years_at_rank,
    certifications,
    area_of_expertise
  ) %>%
  tbl_summary(
    by = domain,
    label = list(
      credentials ~ "Credentials",
      rank ~ "Academic rank",
      years_at_rank ~ "Years at current rank",
      certifications ~ "Certifications",
      area_of_expertise ~ "Area of expertise"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ c(1, 1),
      all_categorical() ~ c(0, 1)
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  modify_header(
    label ~ "**Rater characteristic**",
    stat_0 ~ "**Overall**<br>N = {N}"
  ) %>%
  modify_caption("**Table 1. Characteristics of expert raters by message domain**") %>%
  bold_labels()

table1_raters_by_domain

```

## Rating-Level Summaries

-   For each outcome, report by true source and domain:
-   quality: mean, SD, median, IQR
-   accuracy: mean, SD, median, IQR
-   guideline alignment: n and %
-   perceived source: n and %
-   correct source identification: n and %
-   Also report:
-   total number of ratings
-   average number of ratings per message
-   average number of messages rated per expert

### Quality
```{r}
clean_rating_data

```

### Accuracy
```{r}
clean_rating_data

```

### Guideline Alignment
```{r}
clean_rating_data

```

### Perceived Source
```{r}
clean_rating_data

```

# Primary Research Questions and Models

## Question 1: Do human- and AI-generated messages differ in quality?

```{r}

```

Outcome: quality, 1 to 5.

Use linear mixed-effects regression. Fixed effects:

-   true source

-   domain Random effects:

-   random intercept for message

-   random intercept for expert rater Report:

-   adjusted mean difference for AI versus human

-   95% confidence interval

-   p-value

-   As a sensitivity analysis, we might consider fitting an ordinal mixed-effects model?

## Question 2: Do human- and AI-generated messages differ in accuracy?

Outcome: accuracy, 1 to 5.

Use linear mixed-effects regression. Fixed effects:

-   true source

-   domain Random effects:

-   random intercept for message

-   random intercept for expert rater Report:

-   adjusted mean difference for AI versus human

-   95% confidence interval

-   p-value

-   As a sensitivity analysis, we might consider fitting an ordinal mixed-effects model?

```{r}

```

## Question 3: Do human- and AI-generated messages differ in guideline alignment?

Outcome: guideline alignment, 0 or 1.

Use logistic mixed-effects regression. Fixed effects:

-   true source

-   domain Random effects:

-   random intercept for message

-   random intercept for expert rater Report:

-   odds ratio for AI versus human

-   95% confidence interval

-   p-value

```{r}

```

## Question 4: Can experts correctly identify whether messages were human- or AI-generated?

Analyze source identification in two ways.

### 4a. Three-Category Source Identification Outcome 

Create a three-category source identification variable:

-   correct

-   incorrect

-   not sure

Use multinomial mixed-effects regression, if feasible.

Fixed effects:

-   true source

-   domain

Random effects:

-   random intercept for message

-   random intercept for expert rater

Report:

-   percentage correct, incorrect, and not sure overall

-   percentage correct, incorrect, and not sure by true source

-   percentage correct, incorrect, and not sure by domain

-   model-based estimates for associations between true source, domain, and source-identification category

This should be the primary source-identification analysis because “not sure” was an available response option.

```{r}

```

### 4b. Binary Source Identification Among Forced Classifications 

For this analysis, exclude “not sure” responses. Create a binary source identification variable:

-   1 equals correct

-   0 equals incorrect

Use logistic mixed-effects regression.

Fixed effects:

-   true source

-   domain

Random effects:

-   random intercept for message

-   random intercept for expert rater Report:

-   percentage correct overall

-   percentage correct by true source

-   percentage correct by domain

-   odds ratio

-   95% confidence interval

-   p-value

This binary analysis should be treated as secondary because it excludes “not sure” responses.

```{r}

```

# Session info

```{r session-info}
sessionInfo()
```
