---
title: "AI Project"
format:
  html:
    toc: true
    toc-depth: 5
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
editor: visual
execute:
  echo: true
  include: true
---

Here we explore all our available data files

# Set up

```{r set-up, message = FALSE}
library(tidyverse)
library(readxl)
library(gtsummary)

# library(rio)
# library(lubridate)
# library(stringr)
# library(janitor)
# library(ggplot2)
# library(plotly)
# library(purrr)
# library(reactable)
# library(ggvenn)
# library(naniar)
# library(haven)
```

# Data

## Survey

```{r data}
francois_clean_data <- read_excel("AI_messages_Eval_6-8-26.xlsx", sheet = "Sheet1") 
nutr_rater_data <- read_excel("AI_messages_Eval_6-8-26.xlsx", sheet = "Nutr_Eval_DATA_LABELS_2026-04-0")[, 1:11]
PA_rater_data <- read_excel("AI_messages_Eval_6-8-26.xlsx", sheet = "PA_eval_DATA_Labels")[,1:11]
```

## Basic Check

```{r}
francois_clean_data_basic_check <- francois_clean_data %>%
  mutate(
    check1 = case_when(
      `perceived_source (1=human, 2=AI, 3=not sure)` == "3" ~ NA_real_,
      `true_source (1=human, 2=AI)` == `perceived_source (1=human, 2=AI, 3=not sure)` ~ 1,
      `true_source (1=human, 2=AI)` != `perceived_source (1=human, 2=AI, 3=not sure)` ~ 0
    )
  )

identical(francois_clean_data_basic_check$check1, francois_clean_data_basic_check$source_correct_binary) # TRUE
```

## Clean Survey Data

Rename, recode, and factor some variables for clarity

```{r}
clean_survey_data <- francois_clean_data %>%
  mutate(
    domain = recode(domain, "Physiscal activity" = "physical activity"),
    domain = factor(domain, levels = c("nutrition", "physical activity")),
    true_source = case_when(
      `true_source (1=human, 2=AI)` == 1 ~ "human",
      `true_source (1=human, 2=AI)` == 2 ~ "AI",
      TRUE ~ NA_character_
    ),
    true_source = factor(true_source, levels = c("human", "AI")),
    source_correct_3cat = factor(source_correct_3cat)
  )

str(clean_survey_data)
unique(clean_survey_data$domain)
unique(clean_survey_data$true_source)
```

## Clean Rater Data
```{r}
clean_nutr_rater_data <- nutr_rater_data %>%
  mutate(rater_id = paste0("ne", `Record ID`)) %>%
  dplyr::rename(
    record_id = "Record ID",
    redcap_survey_identifier = "Survey Identifier",
    survey_timestamp = "Survey Timestamp",
    first_name = "First Name",
    last_name = "Last Name",
    credentials = "Highest Degree (e.g. PhD, MD)",
    rank = "Rank (e.g. Postdoc, Instructor, Assistant Professor, etc)",
    years_at_rank = "Years at Rank",
    affiliation = "Institution",
    certifications = "License, Certifications, Professional Certifications (e.g. RDN, ACSM)",
    area_of_expertise = "Area of Research Expertise"
  ) %>%
  select(rater_id, credentials, rank, years_at_rank, affiliation, certifications, area_of_expertise)

clean_PA_rater_data <- PA_rater_data %>%
  mutate(rater_id = paste0("pae", record_id)) %>%
  dplyr::rename(
    survey_timestamp = "evaluation_of_physical_activity_intervention_messa_timestamp",
    first_name = "first_name_v2",
    last_name = "last_name_v2",
    credentials = "credentials_v2",
    affiliation = "affiliation_v2",
    area_of_expertise = "research_expertise"
  ) %>%
  select(rater_id, credentials, rank, years_at_rank, affiliation, certifications, area_of_expertise)

identical(names(clean_nutr_rater_data), names(clean_PA_rater_data))
```


# Descriptive Analyses

## Message-Level Characteristics

-   Report by domain and true source:

-   number of messages

-   number human-generated

-   number AI-generated

```{r}
# Create one-row-per-message dataset
message_level_df <- clean_survey_data %>%
  select(message_id, domain, true_source) %>%
  distinct()

# Table 1: Message characteristics
table1_messages <- message_level_df %>%
  tbl_summary(
    by = true_source,
    include = domain,
    label = list(
      domain ~ "Domain"
    ),
    statistic = list(
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_categorical() ~ c(0, 1),
    missing = "no"
  ) %>%
  add_overall(last = FALSE) %>%
  modify_header(
    label ~ "",
    stat_0 ~ "**Overall**<br>N = {N}",
    stat_1 ~ "**Human-generated**<br>N = {n}",
    stat_2 ~ "**AI-generated**<br>N = {n}"
  ) %>%
  modify_caption("**Table 1. Messages by domain and true source**") %>%
  bold_labels()

table1_messages
```

## Rater-Level Characteristics

-   Report overall and by domain, if relevant:

-   number of raters

-   credentials

-   rank

-   years of rank

-   certifications

-   area of expertise

### Nutrition Rater-Level Characteristics
```{r}
clean_nutr_rater_data

```

### Physical Activity Rater-Level Characteristics
```{r}


```


# Session info

```{r session-info}
sessionInfo()
```
